---
title: "Mo graphics"
author: "Mo Chen"
date: "2022-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###point and line plots
#### Scatterplot
* It is better to separate the dataset from the character in plot
```{r, echo=FALSE}
plot(Petal.Length ~ Petal.Width, data = iris)
```

#### Customization
* Add Lebal
```{r, echo=TRUE}
plot(Petal.Length ~ Petal.Width, data = iris,
     xlab = 'petal.length', 
     ylab = 'petal.width', 
     main = 'iris')
```

* Change symbol type by abbreviation
1. pch = plotting character (by number, number of meaning can be search by Google)
2. cex = character expansion
3. col = (0,0,0,0.1) (the last element is "transparency")

```{r, echo= TRUE}
plot(Petal.Length ~ Petal.Width, data = iris,
     xlab = 'petal.length', 
     ylab = 'petal.width', 
     main = 'iris colorful',
     pch = 19, cex = 2, 
     col = rgb(0,1,0.3,0.1))
```

* different color using
1. if else (with transparency by function "alpha")
```{r, echo= TRUE}
col.iris<-ifelse(iris$Species=='setosa','purple',ifelse(iris$Species=='versicolor','blue','pink')) 
plot(Petal.Length ~ Petal.Width, data = iris,
     xlab = 'petal.length', 
     ylab = 'petal.width', 
     main = 'iris colorful',
     pch = 19, cex = 2, 
     col = scales::alpha(col.iris, 0.3))
```

* discription of plot (Legend)
```{r, echo=TRUE}
col.lev <- as.factor(col.iris)
str(col.lev)
plot(Petal.Length ~ Petal.Width, data = iris,
     xlab = 'Petal width (cm)', 
     ylab = 'Petal length (cm)', 
     main = 'iris with legend',
     cex.axis=1.0, cex.lab=1.5, cex.main=1.5,
     pch = 19, cex=2, las=1,
     col = scales::alpha(col.iris, 0.2))

legend(x="bottomright", pch= 19, cex=1.0,
       legend= c("versicolor","setosa", "virginica"), col=levels(as.factor(scales::alpha(col.lev, 0.8))))
```

### Types
#### panel plot
* scatterplot matrix
```{r, echo=TRUE}
pairs(iris[1:4], pch=19, col = scales::alpha(col.iris, 0.2))
```

* line plot
```{r, echo=TRUE}
blossom <- NULL
blossom$year <- 2010:2019
blossom$count.alaska <- c(3, 1, 5, 2, 3, 8, 4, 7, 6, 9)
blossom$count.canada <- c(4, 6, 5, 7, 10, 8, 10, 11, 15, 17)
as.data.frame(blossom)
plot(count.alaska ~ year,dat = blossom, type='l',
      ylab = "No. of flower blossoming")
```

```{r, echo=TRUE}
y.rng <- range(c(blossom$count.alaska, blossom$count.canada))
plot(count.alaska ~ year,dat = blossom, type='l', ylim = y.rng, pch=20, lwd = 2, col='red', ylab = "No. flower blossoming") 
lines(count.canada ~ year, dat = blossom, type='b', lty = 5, lwd = 1, col='blue')
```

#### Establishing new plot from stratch (null)
```{r, echo=TRUE}
iris.ver<- subset(iris, Species == "versicolor")
iris.vir<- subset(iris, Species == "virginica")

y.rng <- range( c(iris.ver$Petal.Length, iris.vir$Petal.Length) , na.rm = TRUE) 
x.rng <- range( c(iris.ver$Petal.Width, iris.vir$Petal.Width) , na.rm = TRUE) 

# Plot an empty plot

plot(Petal.Length ~ Petal.Width, dat = iris.ver,
     xlab = 'Petal width (cm)', 
     ylab = 'Petal length (cm)', 
     main = 'Petal width and length of iris flower',
     cex.axis=1.0, cex.lab=1.5, cex.main=1.5, type='n',
     xlim=x.rng,  ylim=y.rng)

# Add points for versicolor
points(Petal.Length ~ Petal.Width, dat = iris.ver, pch = 20,cex=2, col = rgb(0,0,1,0.10))
       
# Add points for versicolor
points(Petal.Length ~ Petal.Width, dat = iris.vir, pch = 20,cex=2, 
      col =  scales::alpha('#fc03c6', 0.2))

# Add legend
legend("topleft", c("versicolor", "virginica"), pch = 19, cex=1.2,
       col = c(rgb(0,0,1,0.10), scales::alpha('#fc03c6', 0.2)))
```



