---
title: "Final Report"
author: "Mo Chen, Ethan Wang"
date: "2022-12-27"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The dataset we use is "Drivers of tropical forest loss between 2008 and 2019"
```{r, echo= TRUE}
control <- read.csv("D:/R_WD/Git linked/2022RisFUN-Mo/Final present/Data_fp/ILUC_DARE_controls_x_y.csv")
```

Packages used in this report
```{r, echo= TRUE}
library(ggplot2)
library(ggmap)
library(rnaturalearth)
library(tidyverse)
library(forcats)
library(vegan)
library(RColorBrewer) # color tool
library(ade4) # PCA plot
library(factoextra) # PCA plot
library(magrittr) #PCA plot
library(sp)
library(rworldmap)
```

Data arrangement
```{r, echo= TRUE}
control <- control%>%arrange(answer)
control_pivot <- control%>%pivot_wider(id_cols = c(sampleid,x,y),
                                       names_from = step,values_from = answer,
                                       values_fn = function(x)paste(x,collapse = ","))
control_pivot <- rename(control_pivot, architecture=step3,predominant.driver=step1,other.drivers=step2)
```

Distribution of predominant drivers of global tropical forest loss
```{r, echo= TRUE}
world <- ne_countries(scale = "medium", returnclass = "sf")
predominant.driver <- ggplot(data=world)+
  geom_sf()+
  coord_sf(expand=F)+
  labs(x="Longitude",y="Latitude")+
  geom_point(data=control_pivot,aes(x=x,y=y,colour=factor(predominant.driver)),
             alpha=0.7,size=0.7)+
  ggtitle("Predominant drivers of global tropical forest loss")+
  scale_color_brewer(name="Predominant drivers",palette= "Set1")+
  theme_bw()+
  theme(legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size=7))
predominant.driver
```

Sub plot of regional predominant drivers, separated into America, Asia and Oceania and Africa and Europe
```{r, echo= TRUE}
America.predominant.plot <- ggplot(data=world)+  #south America
  geom_sf()+
  coord_sf(xlim=c(-120,-20),ylim=c(-45,40),expand=F)+
  labs(x="Longitude",y="Latitude")+
  geom_point(data=control_pivot,aes(x=x,y=y,colour=factor(predominant.driver)),
             alpha=0.9)+
  ggtitle("Predominant drivers of forest loss in America")+
  scale_color_brewer(name="Predominant drivers",palette= "Set1")+
  theme_bw()+
  theme(legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size=7))
America.predominant.plot
America <- filter(control_pivot, x<=-20 & x>=-120)
```

```{r, echo= TRUE}
Africa.predominant.plot <- ggplot(data=world)+  #Africa
  geom_sf()+
  coord_sf(xlim=c(-20,55),ylim=c(-35,23.5),expand=F)+
  labs(x="Longitude",y="Latitude")+
  geom_point(data=control_pivot,aes(x=x,y=y,colour=factor(predominant.driver)),
             alpha=0.9)+
  ggtitle("Predominant drivers of forest loss in Africa")+
  scale_color_brewer(name="Predominant drivers",palette= "Set1")+
  theme_bw()+
  theme(legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size=7))
Africa.predominant.plot
Africa <- filter(control_pivot, x<=60 & x>=-20)
```

```{r, echo= TRUE}
Asia.predominant.plot <- ggplot(data=world)+  #Asia
  geom_sf()+
  coord_sf(xlim=c(63,160),ylim=c(-35,40),expand=F)+
  labs(x="Longitude",y="Latitude")+
  geom_point(data=control_pivot,aes(x=x,y=y,colour=factor(predominant.driver)),
             alpha=0.9)+
  ggtitle("Predominant drivers of forest loss in Asia and Oceania")+
  scale_color_brewer(name="Predominant drivers",palette= "Set1")+
  theme_bw()+
  theme(legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size=7))
Asia.predominant.plot
Asia <- filter(control_pivot, x<=160 & x>=63)
```

Ratio boxplot of dirvers in global and regional
```{r, echo= TRUE}
predominant.driver <- fct_count(factor(control_pivot$predominant.driver),sort = T)

predominant.driver$ratio <- predominant.driver$n/sum(predominant.driver$n)

ratio.pre.driver <- ggplot(data = predominant.driver, aes(x=f,y=ratio,fill=f))+
  geom_bar(stat = "identity")+
  scale_fill_manual(name="Predominant drivers",values= c(2:10))+
  geom_text(aes(label=paste(100*round(ratio,3),"%")),vjust=-0.3,size=3)+
  theme_bw()+
  theme(axis.text.x = element_blank())+
  ggtitle("Ratio of predominant tropical forest loss drivers in global")+
  labs(x="Predominant drivers",y="Ratio")
ratio.pre.driver
```

```{r, echo= TRUE}
predominant.drivers.America <- fct_count(factor(America$predominant.driver),sort = T)
predominant.drivers.America$ratio <- predominant.drivers.America$n/sum(predominant.drivers.America$n)
ratio.pre.driver.America <- ggplot(data = predominant.drivers.America, aes(x=f,y=ratio,fill=f))+
  geom_bar(stat = "identity")+
  scale_fill_brewer(name="Predominant drivers",palette = "Set1")+
  geom_text(aes(label=paste(100*round(ratio,3),"%")),vjust=-0.3,size=3)+
  theme_bw()+
  theme(axis.text.x = element_blank())+
  ggtitle("Ratio of predominant tree loss driver in America")+
  labs(x="Predominant drivers",y="Ratio")
ratio.pre.driver.America
```

```{r, echo= TRUE}
predominant.drivers.Africa <- fct_count(factor(Africa$predominant.driver),sort = T)
predominant.drivers.Africa$ratio <- predominant.drivers.Africa$n/sum(predominant.drivers.Africa$n)
ratio.pre.drivers.Africa <- ggplot(data = predominant.drivers.Africa, aes(x=f,y=ratio,fill=f))+
  geom_bar(stat = "identity")+
  scale_fill_brewer(name="Predominant drivers",palette = "Set1")+
  geom_text(aes(label=paste(100*round(ratio,3),"%")),vjust=-0.3,size=3)+
  theme_bw()+
  theme(axis.text.x = element_blank())+
  ggtitle("Ratio of predominant tree loss driver in Africa")+
  labs(x="Predominant drivers",y="Ratio")
ratio.pre.drivers.Africa
```

```{r, echo= TRUE}
predominant.drivers.AO <- fct_count(factor(Asia$predominant.driver),sort = T)
predominant.drivers.AO$ratio <- predominant.drivers.AO$n/sum(predominant.drivers.AO$n)
ratio.pre.drivers.AO <- ggplot(data = predominant.drivers.AO, aes(x=f,y=ratio,fill=f))+
  geom_bar(stat = "identity")+
  scale_fill_brewer(name="Predominant drivers",palette = "Set1")+
  geom_text(aes(label=paste(100*round(ratio,3),"%")),vjust=-0.3,size=3)+
  theme_bw()+
  theme(axis.text.x = element_blank())+
  ggtitle("Ratio of predominant tree loss driver in Asia and Oceania")+
  labs(x="Predominant drivers",y="Ratio")
ratio.pre.drivers.AO
```

PCA plot of predominant drivers in three different regions
```{r, echo= TRUE}
presence.pre.drivers <- control_pivot%>%
 mutate(presence=1)%>%
 pivot_wider(id_cols = c(sampleid,x,y),names_from = predominant.driver,values_from = presence)

presence.pre.drivers[,-(1:3)] <-ifelse(is.na(presence.pre.drivers[,-(1:3)]),0,1)

control_pivot$region <- ifelse(control_pivot$x>=-120 & control_pivot$x<=-20,"America",
                               ifelse(control_pivot$x>=-20 & control_pivot$x<=60,"Africa","Asia and Oceania"))

predominant.pca2 <- dudi.pca(presence.pre.drivers[-(1:3)],scannf=F,nf=6,scale = F) 
fviz_eig(predominant.pca2)
get_eig(predominant.pca2)
fviz_pca_biplot(predominant.pca2,repel = T,habillage=control_pivot$region,
                alpha.ind=0.5, addEllipses = T,ellipse.level=0.95,labelsize=3, title = "PCA - Predominant drivers in three regions")+
  theme(text = element_text(size = 7.5),
        axis.title = element_text(size = 7.5),
        axis.text = element_text(size = 7.5))

```

```{r, echo = TRUE}
countries <- function(points)
{  
  countriesSP <- getMap(resolution='low')
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  
  indices = over(pointsSP, countriesSP)
  indices$ADMIN  #returns country name
}
points <- data.frame(lon=control_pivot$x,lat=control_pivot$y)
control_pivot$country <- countries(points)
presence.pre.drivers <- control_pivot%>%
 mutate(presence=1)%>%
 pivot_wider(id_cols = c(sampleid,x,y,region,country),names_from = predominant.driver,values_from = presence)
presence.pre.drivers[,-(1:5)] <-ifelse(is.na(presence.pre.drivers[,-(1:5)]),0,1)
```

```{r, echo = TRUE}
#沒用國家去分
predominant.pca2 <- dudi.pca(presence.pre.drivers[-(1:5)],
                             scannf=F,nf=6,scale = F)
fviz_eig(predominant.pca2)
get_eig(predominant.pca2)
fviz_pca_biplot(predominant.pca2,repel = T,habillage=control_pivot$region,
                alpha.ind=0.5,addEllipses = T,ellipse.level=0.95,
                labelsize=3)+
  theme(text = element_text(size = 7.5),
        axis.title = element_text(size = 7.5),
        axis.text = element_text(size = 7.5))
```

```{r, echo = TRUE}
#用國家去分
presence.pre.drivers.na.omit <- presence.pre.drivers%>%na.omit(country)
presence.pre.drivers.countries <- presence.pre.drivers.na.omit%>%
  group_by(country)%>%
  summarise(across("Pasture":"Mining and crude oil extraction",.fns=sum))
#percentage
presence.pre.drivers.countries[-1] <- presence.pre.drivers.countries[-1]/rowSums(presence.pre.drivers.countries[-1]) 
countries.continent <- control_pivot%>% #國家跟洲對照表
  select(country,region)%>%
  unique()%>%
  na.omit()%>%
  arrange(country)


predominant.pca4 <- dudi.pca(presence.pre.drivers.countries[,-1],
                             scannf=F,nf=9,scale = F)
fviz_eig(predominant.pca4)
get_eig(predominant.pca4)
fviz_pca_biplot(predominant.pca4,repel = T,
                alpha.ind=0.5,
                labelsize=3,habillage = countries.continent$region,
                addEllipses = T,ellipse.level=0.95)+
  theme(text = element_text(size = 7.5),
        axis.title = element_text(size = 7.5),
        axis.text = element_text(size = 7.5))
```

```{r, echo = FALSE}

```

